---
title: "Pitch_headline_analysis"
author: "M.Kinneen"
date: "2023-01-24"
output: html_document
---
##Set up env
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(fuzzyjoin)
library(data.table)
library(ggplot2)
library(stringr)
library(readr)

#Read in data
sensor_data_processed <- read.csv(file="../data/processed/sensor_data_processed_igfs_2018.csv")
summarised_pitch_per_min <- read.csv(file="../data/processed/pitch_summarised_by_minute.csv")%>%
  mutate(timestamp = as.POSIXct(timestamp))

sensor_data_summary <- read.csv(file="../data/processed/sensor_data_summary_igfs_2018.csv")
poor_weather_hauls<-read.csv("../data/processed/poor_weather_hauls.csv")

```

## Restructure

Create timestamps of the same format for both datasets, which can then be used as primary key.
```{r }

sensor_data_processed%>%
  mutate(SensorTime = NumToTime(SensorTime,as.Date(date_shot)),
         date_shot = as.Date(date_shot))%>%
  mutate(sensor_datetime = 
           as.POSIXct(paste(date_shot,SensorTime),format="%Y-%m-%d %H:%M:%S"
))->sensor_data_processed

```

## Analysis

To explore for a relationship between headline height and vessel pitch, pitch will be observed in 3 different ways.
 i) Summed total absolute pitch per time ( 1,5,10 minute intervals)
 ii) Instantaneous stern velocity (Politis et al)
 iii) 0 crossings?

### Summed absolut pitch per time
```{r }

sensor_data_processed%>%
  select(fldCruiseStationNumber,sensor_datetime,TEY_height,SweepLngt_m)%>%
  group_by(sensor_datetime = as.POSIXct(cut(sensor_datetime, breaks="1 min")),
           fldCruiseStationNumber,SweepLngt_m) %>%
  summarize(trawl_eye_height_median = median(TEY_height,na.rm=TRUE),
            trawl_eye_height_var = var(TEY_height,na.rm=TRUE))%>%
  rename(timestamp = sensor_datetime)->summarised_height_per_min

#Recode

#Join pitch and height datasets
summarised_pitch_per_min<-as.data.table(summarised_pitch_per_min)
summarised_height_per_min<-as.data.table(summarised_height_per_min)

#combine data.tables
pitch_height_combined<-as.data.table(summarised_pitch_per_min)[summarised_height_per_min, on = .(timestamp), roll = TRUE]


#Absolut pitch vs headline per min
#If pattern present, should be observable in hauls with poor weather
#Look at 10 hauls with highest mean wind speed.

#Median headline height
pitch_height_combined%>%
  filter(fldCruiseStationNumber %in% poor_weather_hauls$fldCruiseStationNumber)%>%
  
  ggplot(aes(y=trawl_eye_height_median,x=absolute_pitch_per_min,colour=SweepLngt_m))+
  geom_point()+
  facet_wrap(~fldCruiseStationNumber)


#Median headline height
pitch_height_combined%>%
  filter(fldCruiseStationNumber %in% poor_weather_hauls$fldCruiseStationNumber)%>%
  filter(trawl_eye_height_var<3.5)%>%
  
  ggplot(aes(y=trawl_eye_height_var,x=absolute_pitch_per_min,colour=SweepLngt_m))+
  geom_point()+
  facet_wrap(~fldCruiseStationNumber)




```
```{r}

sensor_data_summary%>%
  filter(ObsPitch_deg < 30 & door_dist_Median > 100)%>%
  
  ggplot(aes(x= ObsPitch_deg,y=door_dist_Median,colour=SweepLngt_m))+
  geom_point()+
  facet_wrap(~SweepLngt_m)+
  stat_smooth(method = "loess")

#For var
sensor_data_summary%>%
  filter(ObsPitch_deg < 30 & door_dist_sd < 10)%>%
  
  ggplot(aes(x= ObsPitch_deg,y=door_dist_sd,colour=SweepLngt_m))+
  geom_point()+
  facet_wrap(~SweepLngt_m)+
  stat_smooth(method = "loess")


library(mgcv)

gam_data<-sensor_data_summary%>%filter(ObsPitch_deg < 30 & door_dist_Median > 70)

gam_1<-gam(TEY_height_sd ~ s(ObsPitch_deg),
    data=gam_data,
    method="REML")

summary(gam_1)
plot(gam_1, residuals = TRUE)


gam_2<-gam(trawl_eye_height_median ~ s(absolute_pitch_per_min),
           data=pitch_height_combined,method="REML")
summary(gam_2)
plot(gam_2)

```




Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
